---
title: "Tidymodels and Vetiver"
format: html
editor_options: 
  chunk_output_type: console
  canonical: true
---

## Goals

The goals of this activity are to:

-   Create two models using the `tidymodels` framework that predict Coca-Cola sales.
-   Save a model as a pin to Posit Connect using the `vetiver` package.
-   Deploy a model to Posit Connect as an API using the `vetiver` package.
-   Learn how to version and monitor your machine-learning models using `vetiver`.

## Overview

1.  Import validated coca-cola sales dataset from the database.
2.  Transform the data for modeling.
3.  Create `logistic_reg()` and `xgboost()` classification models using `tidymodels`.
4.  Assess the performance of both models using `tidymodels`.
5.  Save each model as a pin to Posit Connect using the `vetiver` package.
6.  Use the pinned models to serve a plumber API hosted on Posit Connect.
7.  Use the `vetiver` and `pins` package to version and monitor the performance of the models over time.

## Task 0 - Create and Save a Posit Connect API Key

Before starting this activity, we need to create and save a Posit Connect API key in our environment. This API key identifies you to Connect, allowing you to publish or interact with privileged content if you have the necessary permissions. **YOU SHOULD NEVER ADD YOUR API KEY TO YOUR SOURCE CODE!!!** An API key is like a password, and you should take the necessary steps to secure it. One of those methods is by adding it as an environment variable. You can do this by adding the variable/API key to your `.Renviron` file. This file is interpreted by R every time you start a new session. The `usethis` R packabe makes it super easy to add environment variables.

```{r}
#| eval: false

# Open your .Renviron file
usethis::edit_r_environ()
```

Once you run the above command, you should see a new file open up in RStudio. This is your `.Renviron` file and is usually located in your home (\~) directory. Before we set the variable, we first need to obtain an API key from Connect. Follow the steps below:

1.  Navigate to Posit Connect (TODO: add URL for Connect Server).
2.  Click your username in the top right corner.
3.  Select **API Keys**
4.  Select **+New API Key**
5.  Name it whatever you like. `Connect-API-Key` is a good option.
6.  Copy the new API key. **DON'T LOSE IT!** If you do lose it, you can create another one.

Navigate back to your `.Renviron` file and add the following line. It's always a good idea to add an additional blank line at the end of your `.Renviron` file. Let's also add the CONNECT_SERVER URL just for good measure!

```         
CONNECT_API_KEY=paste-your-api-key-here
CONNECT_SERVER=TODO: Add URL for Connect Server
```

Now you can close the `.Renviron` file and restart your R session (Session–\>Restart R) so that the API key varialble get's added to our environment. To confirm it worked, try running the following code:

```{r}
# Check if CONNECT_API_KEY is added to environment
Sys.getenv("CONNECT_API_KEY")

# Check if CONNECT_SERVER is added to environment
Sys.getenv("CONNECT_SERVER")
```

You should see the API key in the console. If not, please raise your hand and we'll be around to help you out!

## Task 1 - Create a tidymodel

```{r}
#| label: setup

library(tidymodels)
library(dplyr)
library(readr)
library(vetiver)
library(pins)
```

[Tidymodels](https://www.tidymodels.org/) is a collection of R packages for modeling and machine learning using [tidyverse](https://www.tidyverse.org/) principles. While this workshop is **not** a modeling workshop, to demonstrate the utility of the `vetiver`, we need to create a model.

For this activity, we will train and fit a logistic regression model to predict revenue for Coca-Cola sales. Later on, we will create another model using xgboost.

🔄 Tasks

-   Explore the `coca-cola-sales` dataset. Use any method you'd like to understand the data. Try functions like `glimpse()`, `str()`, and `View()`.

-   Create a logistic regression tidymodel that predicts `revenue`.

#### Read and Explore the Data

```{r}
#| label: read-data

# Read in the coca-cola sales data
coca_cola_data <- read_csv("data/coca_cola_sales_data_clean.csv")
```


```{r}
#| label: explore-data

# Explore the data
glimpse(coca_cola_data)
summary(coca_cola_data)
str(coca_cola_data)
View(coca_cola_data)
```

#### Prep Data for Modeling

Sometimes, you may need to transform the original dataset to make it conducive to modeling.

```{r}
#| label: prep-data

model_data <- coca_cola_data |> 
  select(units_sold, unit_price, product, region, store_type, revenue) |> 
  # Convert character variables to factors
  mutate(
    product = as.factor(product),
    region = as.factor(region),
    store_type = as.factor(store_type)
  )
```

#### Split Data

For modeling, you often split your data into a **training and testing** dataset. Before we do that, we'll first extract the last 500 rows, which will be used later on for model monitoring.

```{r}
#| label: split-data

# Split the data
set.seed(123)
data_split <- initial_split(model_data, prop = 0.8, strata = revenue)
train_data <- training(data_split)
test_data <- testing(data_split)

cat("Training set size:", nrow(train_data), "\n")
cat("Test set size:", nrow(test_data), "\n")
```

#### Create the recipe


```{r}
#| label: create-recipe

# Create a simple recipe
revenue_recipe <- recipe(revenue ~ ., data = train_data) |>
  # Create dummy variables for categorical predictors
  step_dummy(all_nominal_predictors()) |>
  # Normalize numeric predictors
  step_normalize(all_numeric_predictors())
```


#### Define model type

```{r}
#| label: define-model-type

# Specify the model (simple linear regression)
linear_model <- linear_reg() |>
  set_engine("lm") |>
  set_mode("regression")
```


#### Create the workflow

```{r}
#| label: create-workflow

revenue_workflow <- workflow() |> 
  add_recipe(revenue_recipe) |> 
  add_model(linear_model)
```

### Train the Model

Train the model using our training set.

```{r}
#| label: train-model

# Train the model
cat("Training model...\n")
fitted_model <- revenue_workflow |> 
  fit(data = train_data)

cat("Model training complete!\n")
```

#### Evaluate the Model

We'll use `tidymodels` to evaluate how our model did using the testing dataset.

```{r}
#| label: evaluate-model

# Make predictions on test set
predictions <- fitted_model |>
  predict(test_data) |>
  bind_cols(test_data)

# Calculate model performance
model_metrics <- predictions |>
  metrics(truth = revenue, estimate = .pred)

cat("\nModel Performance:\n")
print(model_metrics)
```

🛑 Stop here!

## Task 2 - Pin Model to Posit Connect

Now that we have a model, we'll want to save it to a location so that it can be easily shared with colleagues, other content, or your future self! The `pins` package is the workhorse behind this workflow. The `vetiver` package has some wrappers around the pinning functions that we'll use in this task.

🔄 Tasks

-   Create a vetiver model.

-   Pin vetiver model to Posit Connect.

#### Create a Vetiver Model and Pin to Posit Connect

Before we go any further, we need to convert our logistic regression model into a format that can be used by `vetiver`:

```{r}
#| label: create-vetiver-model

# Create a vetiver model
vetiver_model <- vetiver_model(
  model = fitted_model,
  model_name = "coca_cola_revenue_model"
)
```

And now we can save it, or *pin it*, to Posit Connect! Remember to define your pinning board as Posit Connect.

```{r}
#| label: pin-model

# Register Posit Connect as board
board <- board_connect()

# Write Vetiver Model

board |> vetiver_pin_write(vetiver_model)
```

Navigate to Posit Connect and check out your pinned model! You'll also notice a message asking you to create a **Model Card**. This is alsways a good idea, especially if you will be sharing your model with others. An example model card can be found here: `materials/04-tidymodels-vetiver/ferry_modelcard_r.Rmd`.

🛑 Stop here!

## Task 3 - Serve Model as an API

APIs are a great way for users, applications, or other systems to interact with your model. `vetiver` leverages [Plumber](https://www.rplumber.io/) which is used to create APIs with only R code! Let's create an API here within Posit Workbench, and then deploy it to Posit Connect.

🔄 Tasks

-   Deploy model to Posit Connect as an API.

-   Make a prediction using the API visual interface (RapiDoc).

-   Make a request to the API using R code.

#### Deploy API to Posit Connect

**Be sure to add your user name for the pin!**

**Be sure to add your user name for the pin!**

```{r}
#| label: create-api

# Create API from pinned model using vetiver
#  Swap out "____" for your username on Connect.
vetiver_deploy_rsconnect(board = board, name = "ryan/coca_cola_revenue_model")
```

Once it's deployed, navigate to the API on Posit Connect and give it a test run! Click the operation on the left called `Return predictions from model using 5 features`. Then copy any variation of the request below and click **Try**. You should see a response of the predicted revenue:

```
[
  {
    "units_sold": 150,
    "unit_price": 2.00,
    "product": "Sprite",
    "region": "Europe",
    "store_type": "Supermarket"
  }
]
```

#### Make a Prediction using R Code

On Posit Connect, navigate to your newly created API, and copy the URL for the content below. You can find this URL by clicking the access tab --\> URL.

```{r}
#| label: create-endpoint

api_url <- "https://connect.posit.it/content/d38c29b8-250f-4a22-ba05-4ff4a49f0304"

# Append "/predict" to the end of your api to create the endpoint
endpoint <- vetiver_endpoint(paste0(api_url, "/predict"))

endpoint
```

Below is a data frame representing information for a Coca-Cola sale. Let's see what the predicted revenue will be. Feel free to modify.

```{r}
#| label: make-new-coke-sale-data

# Create new ferry data point
new_sale <- tibble(
  units_sold = 300,
  unit_price = 1.50,
  product = "Sprite",
  region = "Europe",
  store_type = "Supermarket"
)
```

Now we can use the `predict()` function to query the API so that it returns a revenue prediction. Note that we have to add our API key to the predict function since this API is only accessible to specific users!

```{r}
#| label: predict-r

# Predict from Vetiver endpoint
predict(endpoint, 
        new_sale,
        httr::add_headers(Authorization = paste("Key", Sys.getenv("CONNECT_API_KEY"))))$.pred
```

🛑 Stop here!