---
title: "Validate Cola Sales Data"
format:
  html:
    toc: true
    toc-location: left
    anchor-sections: true
    code-fold: show
    code-overflow: wrap
    code-summary: "Show Code"
    code-tools: true
    code-link: true
execute:
    eval: false
editor_options: 
  chunk_output_type: console
  canonical: true
html-table-processing: none
---

## Goals

The goals of this activity are to:

-   Use the `pointblank` package to perform data validations

## Setup

```{r}
#| label: setup

library(paws)
library(tidyverse)
library(pointblank)
```

Load the cola sales data.

```{r}
#| label: load-data

# Make a client service to interact with S3
s3_client <- paws::s3()

# Bucket information
bucket_name <- "ptd-demo-bucket"
file_key <- "posit_cola_sales_data.csv"       

# Read the data from S3
response <- s3_client$get_object(
  Bucket = bucket_name,
  Key = file_key
)

data <- read_csv(rawToChar(response$Body))

```

We will want to verify our column schema matches what is expected. Let's define our schema for use in our validation plan.

```{r}
# Two helpful functions for understanding the column types
data |> glimpse()

data |> spec()
```

```{r}
# define our column schema, which we will use in validation
schema_data <- col_schema(
  date = "Date",
  product = "character", 
  region = "character",
  store_type = "character",
  store_id = "numeric",
  units_sold = "numeric",
  unit_price = "numeric",
  revenue = "numeric",
  promotion = "character"
)
```

## Task 1 - Create a validation plan

To create a validation plan using `pointblank` we need to:

-   create an agent
-   define validations

```{r}
#| label: create agent for posit cola data

agent <- pointblank::create_agent(____)
agent
```

On its own, the agent is not very informative. It's waiting for validations to be defined and an interrogation action to be performed.

Now we define our **data validation functions**. A few have been started for you as examples. It's up to you to fill in the suggested validations. Refer to the package documentation for the validation function reference: <https://rstudio.github.io/pointblank/reference/index.html#validation-expectation-and-test-functions>

```{r}
#| label: define validations for posit cola sales data

agent <- create_agent(data, label = "Cola Sales Data Validation") |> 
  # Verify schema
  col_schema_match(schema_data) |>
  # Required fields are present
  col_vals_not_null(columns = c(date, product, region, store_id, units_sold, unit_price)) |>
  # Ensure all rows are distinct (no dupliates)
  ____ |>
  # Validate that `units_sold`, `unit_price`, and `revenue` are reasonable values
  ____ |>
  ____ |>
  ____ |>
  # Validate product names are among the expected set of names
  ____ |>
  # Validate date is in expected range
  ____ |>
  # What else would you validate?
  ____ |> 
  ____ 


agent
```

Tip: Use the Positron Assistant to generate the additional validations:

> Create pointblank validations where there are ___ placeholders.

If we look at the output of `agent`, it shows our validation plan, but the action is yet to come. Next step is to run it with **interrogate**!

## Task 2 - Run validation plan

We use the `interrogate` function to run agent through the validation plan.

```{r}
#| label: interrogate 

agent <- agent |> interrogate()

agent
```

Explore the validation report. Can you:

1.  Identify what fail percentage each validation had?
2.  Identify how many rows failed each validation?
3.  Notice the left-hand side color bands to indicate pass condition?

Change the parameters of your validations to trigger more failures just to see the consequence.

## Task 3 - Remove failing data from the data set

Pointblank has identified all of the rows of `data` that passed and failed validation. Now remove those that failed so the data that is passed downstream to our modeling step is squeaky clean.

Pointblank provides a number of [post-interrogation functions](https://rstudio.github.io/pointblank/reference/index.html#post-interrogation) to work with intel gathered from the validation. For this task, we will "sunder" the data using `pointblank::get_sundered_data()`.

> **💡 sunder** /sun·der / ˈsən-dər / *verb* \| to split apart

```{r}
#| label: sunder data

# Passed data
data_validated <- get_sundered_data(agent = ____, type = "____")

# Failed data
data_faulty <- get_sundered_data(agent = ____, type = "____")

```

## Task 4 - Post-interrogation logicals

Pointblank interrogation provides multiple layers of information about our data. We can take advantage of this with logical TRUE / FALSE statements that drive downstream tasks or effects.

-   Use `pointblank::all_passed()` to determine if all validation steps passed

-   Use `pointblank::get_agent_x_list` to determine if any warnings were set

```{r}
#| label: All validations passed?

# Did all validations pass?
pointblank::____

```

A broad all passed / failed for the entire validation plan might not provide enough granularity for a downstream task. We can drill into more details about each step of the validation and the results using the agent "x_list".

First we will see what the x_list contains.

```{r}
#| label: get agent x list

xlist <- pointblank::get_agent_x_list(agent)

xlist
```

The output is like a gold mine! The resulting list includes the pass / fail statistics (e.g., `$n_passed`, `$n_failed`, `$f_passed`, `$f_failed`). Take a look:

```{r}
#| label: extract the pass / fail statistics

xlist$n_passed

xlist$n_failed

xlist$____
```

The xlist also includes which records triggered a threshold for `warn`, `stop`, or `notify` -- things we will talk about next when we talk about alerting!