---
title: "Validate Cola Sales Data (with conditional alerting)"
format:
  email:
    toc: true
    toc-location: left
    anchor-sections: true
    code-fold: show
    code-overflow: wrap
    code-summary: "Show Code"
    code-tools: true
    code-link: true
editor: visual
editor_options: 
  chunk_output_type: console
  canonical: true
html-table-processing: none
---

## Goals

The goals of this activity are to:

-   Iterate on our previous validation plan to include thresholds for alerting
-   Trigger different conditional emails based on the validation thresholds

## Setup

```{r}
#| label: setup

library(paws)
library(tidyverse)
library(pointblank)
library(quarto)
library(glue) 
library(gt)

```

Load the cola sales data.

```{r}
#| label: load-data

# Make a client service to interact with S3
s3_client <- paws::s3()

# Bucket information
bucket_name <- "ptd-demo-bucket"
file_key <- "posit_cola_sales_data.csv"       

# Read the data from S3
response <- s3_client$get_object(
Bucket = bucket_name,
Key = file_key
)

data <- read_csv(rawToChar(response$Body))

```

We will want to verify our column schema matches what is expected. Let's define our schema for use in our validation plan.

```{r}
# define our column schema, which we will use in validation
schema_data <- col_schema(
  date = "Date",
  product = "character", 
  region = "character",
  store_type = "character",
  store_id = "numeric",
  units_sold = "numeric",
  unit_price = "numeric",
  revenue = "numeric",
  promotion = "character"
)
```

## Task 1 - Create a validation plan with thresholds for alerting

Iterate on the agent we created in the previous exercise to add action levels. Action levels behave like tags. You can decide what threshold you want to put for `notify`, `warn`, and `stop`. At a minimum, the tag will provide a visual indicator of threshold exceedance in the validation table. You can also use these tags post-interrogation to take actions.

The action levels can be set as an **integer**, representing the threshold number of failing units, or a **fraction**, representing the fraction of failing units.

Use `actions = action_levels(warn_at = ____, stop_at = ____, notify_at = ____)` to add desired action levels to one, some, or all of your validations and rerun the interrogation to see the output. Some samples have been provided.

```{r}
#| label: define validations for posit cola sales data

agent <- create_agent(data, label = "Cola Sales Data Validation") |>
  # Verify schema
  col_schema_match(
    schema_data,
    # ðŸ‘‰ use 1 failure as a threshold for a stop flag
    actions = action_levels(stop_at = 1)
  ) |>
  # Required fields are present
  col_vals_not_null(
    columns = c(date, product, region, store_id, units_sold, unit_price),
    # ðŸ‘‰ set a warn flag if more than 1% of the data has missing values;
    # ðŸ‘‰ set a stop flag if more than 10% of the data has missing values
    actions = action_levels(warn_at = 0.01, stop_at = 0.10)
  ) |>
  # Ensure all rows are distinct (no dupliates)
  rows_distinct() |>
  # Validate that `units_sold`, `unit_price`, and `revenue` are reasonable values
  col_vals_gt(columns = units_sold, value = 0) |>
  col_vals_gt(columns = unit_price, value = 0) |>
  col_vals_gt(columns = revenue, value = 0) |>
  # Validate product names are among the expected set of names
  col_vals_in_set(
    columns = product,
    set = c("Posit Classic", "Posit Lite", "Workbench Energy", "Quarto Fizz", "Connect Orange", "Posit Zero", "Cloud Water"),
    # ðŸ‘‰ Try narrowing the set of allowed product names to inflate the number of failures, then set different actions by uncommenting the line below
    # actions = action_levels(warn_at = ____)
  ) |>
  # Validate date is in expected range
  col_vals_between(
    columns = date,
    left = as.Date("2021-01-01"),
    right = as.Date("2024-12-31")
  ) |>
  # interrogate
  interrogate()


agent

```

## Task 2 - Were any action levels set?

Remember the xlist? The information about whether any of our action flags were set.

```{r}
#| label: get agent x list

xlist <- pointblank::get_agent_x_list(agent)

xlist

```

```{r}

stop_notice <- any(xlist$stop)
warn_notice <- any(xlist$warn)

```

## Task 3 - Set conditionals depending on validation outcome and write Quarto metadata

We want the results of our validation to trigger different alerts and actions. Namely:

a.  If no action levels were set, save the validated data for the model to use, but don't send an email. No reason to clutter the inbox!
b.  If a warn level was set, still save our validated data for the model to use, but send an email that something might require a closer look
c.  If a stop level was set, do not save the data for downstream use, and send an alert email

First use some logic to determine what condition we are in:

```{r}
#| label: set conditionals

if (isTRUE(stop_notice)){
  condition <- "STOP"
  send_email <- TRUE
} else if (isTRUE(warn_notice)){
  condition <- "WARN"
  send_email <- TRUE
} else {
  condition <- "OKAY"
  send_email <- FALSE
}

glue::glue("condition set to: ", {condition})
glue::glue("send_email set to: ", {send_email})

```

We will select which email gets sent by writing metadata to Quarto. The `quarto` R package includes a function `write_yaml_metadata_block` which will allow dynamic creation of metadata that is read at document render.

Importantly, in the code block, you must include `#| output: asis` for correct rendering.

```{r}
#| label: metadata-block
#| output: asis

quarto::write_yaml_metadata_block(
  # the yaml will be of the form `is_email_variant_{condition}: true`
  # using setNames to dynamically set the key name
  .list = setNames( 
    nm = glue::glue("is_email_variant_{condition}"), list(TRUE)
    )
)


# this will output yaml-formmatted text into the Quarto document, interpreted at render. Example:
#---
#is_email_variant_WARN: true
#---


```

## Task 4 - Conditionally write the validated data for downstream use

As long as we don't have a STOP condition, we will write the validated data to a pin for the model to use in our next steps for this project.

```{r}
#| label: sunder data

# sunder the data (aka remove the failed records)
data_validated <- get_sundered_data(agent = agent, type = "pass")

```

```{r}
#| label: write data
#| eval: false 

# note, for the workshop, we set `eval: false` for this chunk because participants do not have write access to the S3 bucket

# Conditionally write back to S3

if (condition %in% c("WARN","OKAY")){
    
    file_key <- "posit_cola_sales_data_validated.csv"
    
    # Convert the dataframe to CSV format
    csv_data <- readr::format_csv(data_validated)
    
    # Write the CSV data to S3
    s3_client$put_object(
      Bucket = bucket_name,
      Key = file_key,
      Body = csv_data
    )
    
}

```

## Task 5 - Write our email

Use Quarto fenced divs to define an email with different bodies depending on the metadata.

And if `send_email` is false (because of `condition == OKAY`), no email will be sent.

::::::: email
::: email-scheduled
```{r}
send_email 
```
:::

::: subject
```{r}
#| label: define subject
#| echo: false
#| output: asis


if (condition == "STOP"){
  glue::glue("âš ï¸ ALERT: Posit Cola sales data integrity issue")
} else if (condition == "WARN"){
  glue::glue("â„¹ï¸ Heads Up: Posit Cola sales data updated with warnings")
} else {
  NA
}


```
:::

::: {.content-visible when-meta="is_email_variant_STOP"}
## âš ï¸ There was a data integrity issue with the project data on `{r} today()`

The Posit Cola sales data for `{r} today()` **failed** schema verification, indicating a data integrity issue.

Further processing of the data did not take place. **Downstream data and artifacts have not been updated.**

Please review the raw data to diagnose the issue. An extract of the raw data is shown below:

```{r}
#| echo: false

head(data)

```
:::

::: {.content-visible when-meta="is_email_variant_WARN"}
## â„¹ï¸ Posit Cola sales data updated for `{r} today()` with warnings

The Posit Cola sales data for modeling has been validated and written, but there were warnings set due to exceedance of validation thresholds.

### Validation failures

The table below summarizes the validation step(s) that triggered warnings.

```{r}
#| label: get failing validations details
#| echo: false

if (condition == "WARN") {
  step <- xlist$i[which(xlist$warn)]
  desc <- xlist$briefs[which(xlist$warn)]
  fails <- xlist$f_failed[which(xlist$warn)]
  threshold <- purrr::map(step, ~{xlist$validation_set$actions[[.x]]$warn_fraction}) |> unlist()

  warnings_table <- data.frame(step, desc,threshold,fails) |> gt() |> cols_label(step="Validation Step",desc="Description",fails="Fraction Failed", threshold="Threshold") |> fmt_number(columns = c(fails, threshold),decimal = 2)

  warnings_table |> as_raw_html() 
}
```

Review the associated extract files (attached) to see the records that failed validation.

```{r}
#| label: get extracts and attach to email
#| output: asis
#| echo: false

if (condition == "WARN") {
  failures <- purrr::map(
    step,
    ~ {
      get_data_extracts(agent, .x)
    }
  )

  purrr::walk2(
    failures,
    step,
    ~ {
      write_csv(.x, file = glue::glue("extracts_step_{.y}.csv"))
    }
  )

  filenames_yaml <- glue::glue(" - extracts_step_{step}.csv") |>
    str_c(collapse = "\n")

  quarto::write_yaml_metadata_block(
    .list = list(
      `email-attachments` = filenames_yaml
    )
  )
}
```
:::
:::::::

## Bonus - Add logging to this report

**Report run** `{r} today()`

**Condition set:** `{r} condition`

**Summary of results:**

```{r}
#| label: logging summary
#| echo: false
#| output: asis

if (condition == "STOP") {
  glue::glue("Data integrity validation failed. Data processing for the model training did not proceed. An email notification of the failure was sent.")
}

if (condition == "WARN") {
  glue::glue("Posit Cola sales data was validated and sundered, however a warning was triggered due to one or more validation steps exceeding the threshold for allowed failures.

Records removed that failed validation: {nrow(data) - nrow(data_validated)}

Records written to database: {nrow(data_validated)}

Warnings triggered for validation step(s) listed below:")

  warnings_table
}

if (condition == "OKAY") {
  glue::glue("Posit Cola sales data was validated and sundered.

Records removed that failed validation: {nrow(data) - nrow(data_validated)}

Records written to database: {nrow(data_validated)}")
}

```

## Preview of email sent:

<iframe src="email-preview/index.html" width="100%" height="500">

</iframe>